---
title: "P8105_hw2_jg4890"
author: "Jiayi Ge"
date: "2024-09-30"
output: github_document
---
```{r}
library(tidyverse)
```

## Import data for Q1
```{r}
nyc_data <- read_csv("./nyc_data.csv")
```

## Read and clean the data for Q1
```{r}
nyc_clean <- nyc_data %>%
  select(Line, `Station Name`, `Station Latitude`, `Station Longitude`, Route1, Route2, Route3, Route4, Route5, Route6, Route7, Route8, Route9, Route10, Route11, `Entrance Type`, Entry, Vending, ADA)

nyc_clean <- nyc_clean %>%
  mutate(Entry = ifelse(Entry == "YES", TRUE, FALSE))

dim(nyc_clean) 
# Display (rows x columns) of the cleaned dataset

print(nyc_clean)
```
This orginal dataset contains key variables including, Division, Line, Station Name, Station Latitude and Station Longitude, Routes (1-11), Entrance Type(stair, elevator, easement or door), Exit Only, Entry (whether entry is allowed), Vending (whether vending machines are present), Entrance Type (the type of entrance or exit), and ADA (whether the station is ADA compliant).

In the data cleaning process, several columns were retained, including line, station name, latitude/longitude, routes, entry, vending, entrance type, and ADA compliance. The Entry variable, which was initially in character form ("YES" or "NO"), was converted to a logical variable (TRUE or FALSE) for easier analysis. The dataset was checked for missing values, and unnecessary columns were removed.

The resulting cleaned dataset has dimensions of 1868 rows and 19 columns, and it is considered tidy. Each variable represents a single feature, and each row represents a unique station entrance or exit, ensuring the data follows the principles of tidy data.

**Answer the question in Q1**
```{r}
distinct_stations <- nyc_clean %>%
  distinct(`Station Name`, Line) %>%
  nrow()
distinct_stations

ada_compliant_stations <- nyc_clean %>%
  filter(ADA == TRUE) %>%
  distinct(`Station Name`, Line) %>%
  nrow()

ada_compliant_stations

proportion_allow_entrance <- nyc_clean %>%
  filter(Vending == "NO") %>%
  summarise(proportion = mean(Entry == TRUE))

proportion_allow_entrance
```
There are 465 distinct stations, and 84 stations are ADA compliant. 0.377 of station entrances / exits without vending allow entrance.

## Stations serve the A train
```{r}
library(dplyr)
distinct_a_train_stations <- nyc_clean %>%
  filter(if_any(starts_with("Route"), ~ . == "A")) %>%  
  distinct(`Station Name`, Line) %>% 
  nrow()                                      

distinct_a_train_stations

ada_compliant_a_train_stations <- nyc_clean %>%
  filter(if_any(starts_with("Route"), ~ . == "A")) %>%
  filter(ADA == "TRUE") %>%
  distinct(`Station Name`, Line) %>%
  nrow()

ada_compliant_a_train_stations
```
60 distinct stations serve the A train, and 17 of them are ADA compliant.

## Question 2 ##
Import data and cleaning data
```{r}
library(readxl)
library(dplyr)
mr_trash_wheel <- read_excel("./Mr. Trash Wheel Collection Data.xlsx", sheet = "Mr. Trash Wheel")

mr_trash_wheel_cleaned <- mr_trash_wheel %>%
  select(-c(`...15`, `...16`))

mr_trash_wheel_cleaned <- mr_trash_wheel_cleaned %>%
  filter(!is.na(Dumpster))%>% 
  rename(
    weight_tons = `Weight (tons)`, 
    volume_cubicyeards = `Volume (cubic yards)`,
    sports_balls = `Sports Balls`)%>%
  mutate(ports_balls = as.integer(round(sports_balls)), 
        )
  
head(mr_trash_wheel_cleaned)  

```


Import, clean, and organize the data for Professor Trash Wheel
```{r}
prof_trash_wheel <- read_excel("./Mr. Trash Wheel Collection Data.xlsx", sheet = "Professor Trash Wheel")

prof_trash_wheel_cleaned <- prof_trash_wheel %>%
  filter(!is.na(Dumpster))%>% 
  rename(
    weight_tons = `Weight (tons)`, 
    volume_cubicyeards = `Volume (cubic yards)`)%>%
    mutate(Year = as.character(Year))
head(prof_trash_wheel_cleaned)  

```

Import, clean, and organize the data for Gwynnda Trash Wheel
```{r}
gwy_trash_wheel <- read_excel("./Mr. Trash Wheel Collection Data.xlsx", sheet = "Gwynnda Trash Wheel")

gwy_trash_wheel_cleaned <- gwy_trash_wheel %>%
  filter(!is.na(Dumpster))%>% 
  rename(
    weight_tons = `Weight (tons)`, 
    volume_cubicyeards = `Volume (cubic yards)`) %>%
    mutate(Year = as.character(Year))
head(gwy_trash_wheel_cleaned)  
```
The `Year` was found not in the same data type as the Mr. trash wheel one, so I convert the prof and gwynnda ones into the character.

