---
title: "P8105_hw2_jg4890"
author: "Jiayi Ge"
date: "2024-09-30"
output: github_document
---
```{r}
library(tidyverse)
library(janitor)
```

## Import data for Q1
```{r}
nyc_data <- read_csv("./nyc_data.csv")
```

## Read and clean the data for Q1
```{r}
nyc_clean <- nyc_data %>%
  select(Line, `Station Name`, `Station Latitude`, `Station Longitude`, Route1, Route2, Route3, Route4, Route5, Route6, Route7, Route8, Route9, Route10, Route11, `Entrance Type`, Entry, Vending, ADA)

nyc_clean <- nyc_clean %>%
  mutate(Entry = ifelse(Entry == "YES", TRUE, FALSE))

dim(nyc_clean) 
# Display (rows x columns) of the cleaned dataset

print(nyc_clean)
```
This orginal dataset contains key variables including, Division, Line, Station Name, Station Latitude and Station Longitude, Routes (1-11), Entrance Type(stair, elevator, easement or door), Exit Only, Entry (whether entry is allowed), Vending (whether vending machines are present), Entrance Type (the type of entrance or exit), and ADA (whether the station is ADA compliant).

In the data cleaning process, several columns were retained, including line, station name, latitude/longitude, routes, entry, vending, entrance type, and ADA compliance. The Entry variable, which was initially in character form ("YES" or "NO"), was converted to a logical variable (TRUE or FALSE) for easier analysis. The dataset was checked for missing values, and unnecessary columns were removed.

The resulting cleaned dataset has dimensions of `r nrow(nyc_clean)` rows and `r ncol(nyc_clean)` columns, and it is considered tidy. Each variable represents a single feature, and each row represents a unique station entrance or exit, ensuring the data follows the principles of tidy data.

**Answer the question in Q1**
```{r}
distinct_stations <- nyc_clean %>%
  distinct(`Station Name`, Line) %>%
  nrow()
distinct_stations

ada_compliant_stations <- nyc_clean %>%
  filter(ADA == TRUE) %>%
  distinct(`Station Name`, Line) %>%
  nrow()

ada_compliant_stations

proportion_allow_entrance <- nyc_clean %>%
  filter(Vending == "NO") %>%
  summarise(proportion = mean(Entry == TRUE))

proportion_allow_entrance
```
There are 465 distinct stations, and 84 stations are ADA compliant. 0.377 of station entrances / exits without vending allow entrance.

## Stations serve the A train
```{r}
library(dplyr)
distinct_a_train_stations <- nyc_clean %>%
  filter(if_any(starts_with("Route"), ~ . == "A")) %>%  
  distinct(`Station Name`, Line) %>% 
  nrow()                                      

distinct_a_train_stations

ada_compliant_a_train_stations <- nyc_clean %>%
  filter(if_any(starts_with("Route"), ~ . == "A")) %>%
  filter(ADA == "TRUE") %>%
  distinct(`Station Name`, Line) %>%
  nrow()

ada_compliant_a_train_stations
```
60 distinct stations serve the A train, and 17 of them are ADA compliant.

## Question 2 ##
Import data and cleaning data
```{r}
library(readxl)
mr_trash_wheel <- read_excel("./Trash Wheel Collection Data.xlsx", sheet = "Mr. Trash Wheel")

#Since there are two rows without any info, I delect them.
mr_trash_wheel_cleaned <- mr_trash_wheel %>%
  janitor::clean_names()

mr_trash_wheel_cleaned <- mr_trash_wheel_cleaned %>%
  filter(!is.na(dumpster))%>%
  mutate(sports_balls = as.integer(round(sports_balls)), 
        )
  
head(mr_trash_wheel_cleaned)  

```


Import, clean, and organize the data for Professor Trash Wheel
```{r}
prof_trash_wheel <- read_excel("./Trash Wheel Collection Data.xlsx", sheet = "Professor Trash Wheel")%>%
  janitor::clean_names()
prof_trash_wheel_cleaned <- prof_trash_wheel %>%
  filter(!is.na(dumpster))%>% 
    mutate(year = as.character(year))
  
prof_trash_wheel_cleaned 

```

Import, clean, and organize the data for Gwynnda Trash Wheel
```{r}
gwy_trash_wheel <- read_excel("./Trash Wheel Collection Data.xlsx", sheet = "Gwynnda Trash Wheel")%>%
  janitor::clean_names()

gwy_trash_wheel_cleaned <- gwy_trash_wheel %>%
  filter(!is.na(dumpster))%>% 
    mutate(year = as.character(year))
gwy_trash_wheel_cleaned  
```
The `Year` was found not in the same data type as the Mr. trash wheel one, so I convert the prof and gwynnda ones into the character.

**Produce a single tidy dataset**
```{r}
mr_trash_wheel_clean <- mr_trash_wheel_cleaned %>%
  mutate(Trash_Wheel = "Mr. Trash Wheel")

prof_trash_wheel_clean <- prof_trash_wheel_cleaned %>%
  mutate(Trash_Wheel = "Professor Trash Wheel")

gwynnda_trash_wheel_clean <- gwy_trash_wheel_cleaned %>%
  mutate(Trash_Wheel = "Gwynnda")

all_trash_wheels <- bind_rows(mr_trash_wheel_clean, prof_trash_wheel_clean, gwynnda_trash_wheel_clean)

glimpse(all_trash_wheels)
```
The dataset contains information collected by three trash wheels: Mr. Trash Wheel, Professor Trash Wheel, and Gwynnda. Key variables in the dataset include the date of collection, the volume and weight of trash collected, the number of cigarette butts, sports balls, and other trash-related data. The dataset contains `r nrow(all_trash_wheels)` rows and `r ncol(all_trash_wheels)` columns.


Calculate total weight of trash collected by Professor Trash Wheel
```{r}
prof_trash_weight <- all_trash_wheels %>%
  filter(Trash_Wheel == "Professor Trash Wheel") %>%
  summarize(total_weight = sum(weight_tons, na.rm = TRUE))
prof_trash_weight
```
Total weight is `r prof_trash_weight` tons.


Calculate total number of cigarette butts collected by Gwynnda in June of 2022
```{r}
gwynnda_cig_butts_june2022 <- all_trash_wheels %>%
  filter(Trash_Wheel == "Gwynnda", 
         format(date, "%Y-%m") == "2022-06") %>%
  summarize(total_cig_butts = sum(`cigarette_butts`, na.rm = TRUE))

gwynnda_cig_butts_june2022
```
Total number of cigarette butts collected by Gwynnda in June of 2022 is `r gwynnda_cig_butts_june2022`


## Question 3

Import Data, Data Cleaning and Tidying
```{r}
library(tidyverse)

bakers <- read_csv("./gbb_datasets/bakers.csv")
bakes <- read_csv("./gbb_datasets/bakes.csv")
results <- read_csv("./gbb_datasets/results.csv", skip = 2)
#The very first two rows for "results" is not useful and will make the code can't run correctly.
viewers <- read_csv("./gbb_datasets/viewers.csv")

bakers_clean <- bakers %>%
  janitor::clean_names()

bakes_clean <- bakes %>%
  janitor::clean_names()

results_clean <- results %>%
  janitor::clean_names()

colnames(bakers_clean)
colnames(bakes_clean)
colnames(results_clean)

bakers_clean <- bakers %>%
  mutate(Series = as.numeric(Series), 
         `Baker Age` = as.numeric(`Baker Age`)) 

bakes_clean <- bakes %>%
  left_join(bakers, by = c("Baker" = "Baker Name", "Series" = "Series")) 
  
results <- results %>%
  rename(Baker = baker, Series = series, Episode = episode)
  
merged_data <- bakes %>%
  left_join(bakers_clean, by = c("Baker" = "Baker Name", "Series" = "Series")) %>%
  left_join(results, by = c("Baker" = "Baker", "Series" = "Series", "Episode" = "Episode"))

missing_data <- merged_data %>%
  summarise_all(~ sum(is.na(.)))

final_data <- merged_data %>%
  select(-`Baker Age`, -`Baker Occupation`, -`Hometown`) %>%
  arrange(Series, Episode, Baker)

write_csv(final_data, "./gbb_datasets/final_data.csv")

# During doing this csv file, the most important thing is to know which is the data we need and try to collect them. Before doing all the programming, I observed these files first, to know the key variables. When I try to combine the "result" file, I met some problems due to the very first two empty rows. After skip the unnecessary data, I changed all the names into the same ones, which make these files can merge correctly. Also, after ran the final file, I deleted the useless columns. The final data I had, is combined with the most useful parts, including the Series, Episode, Baker, Signature Bake, Show Stopper, technical and result.
```

Answer questions for Q3
```{r}
star_baker <- final_data %>%
  filter(Series >= 5 & Series <= 10, result == "STAR BAKER" | result == "WINNER") %>%
  select(Series, Episode, Baker, result)

star_baker %>%
  arrange(Series, Episode, Baker, result)
```

```{r}
library(tidyr)
library(readr)
viewers <- read_csv("./gbb_datasets/viewers.csv")
print(head(viewers, 10))

avg_Series_1 <- mean(viewers$`Series 1`, na.rm = TRUE)

avg_Series_5 <- mean(viewers$`Series 5`, na.rm = TRUE)

print(avg_Series_1)
print(avg_Series_5)
```
The average viewership in Season 1 is 2.77.  In Season 5 is 10.0. 

